---
title: "Dogecoin to the ground"
subtitle: "A forecasting of the Dogecoin price "
author: Siming Shan
toc: TRUE
abstract: 
output:
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction:
Elon Musk is known as one of the richest men in the world, and the CEO of the Tesla Motors. On March 14th, Elon Musk proposed to buy the whole twitter company with price of 43 billion dollars has brought him to the public attention again, however,as a matter of fact, Elon Musk has bought 9% stake of twitter and this news caused the twitter shares soared more than 20% []. It is not the first time Elon Musk has had a huge impact on the finance market.

last year was a big year of crypto currency because Elon Musk greatly advertise a cryto currency that he bought - Dogecoin. Dogecoin is a open-source cryptocurrency created by Billy Markus and Jackson Palmer[]. The cryptocurrency is a digital or virtual currency which is secured by crypotography[]. In 2018, as Satoshi Nakamoto created the first cryptocurrency bitcoin in the world, the concept of decentralized networks transaction has been more and more accepted and populated around the world. Nowasays, the crypotocurrency investment has become the first choice of many investors, because it is secure, decentralized, and stable.

On Elon Musk twitter, he claimed that the "dogecoin is the people's crypto", and on May 10th 2021, he claimed that SpaceX will launch satellite called "Doge-1" to the moon, thus "Doge to the moon" has become a famous meme in the crypto currency area. As a matter of fact, the doge coin did actually soar up after Elon Musk's tweets. However, after the tide of dogecoin has passed, and Elon Musk decreased his frequency of tweeting about dogecoin, the price of dogecoin rapidly fell. Does the dogecoin still worthy to invest? This paper uses the statistical programing language R[] and fucous on using the auto-regressive integrated moving average(ARIMA) model to modeling the dogecoin price from 1st May 2021 to 23rd April 2022 to forecast the dogecion price in the next ten day. The data was collected from Yahoo finance by using the function `getSymbols` from the `quantmod` package[]. In the orginal dataset, there are 358 observations with six variables each, these variables are the open price, the high price, the low price, the close price, the volume, and the adjusted price. We focus on the close price since it represents the last price during a regular trading day.

ARIMA is a statistical anlysis model used to forecast time series dataset such as the unemplyment rate, stock price, or the global temperature. The ARIMA can forest a future value Based on the past values and past errors, the mathematical equation of this model can be expressed as follows:
\[X_t = \phi_0 + \phi_1 X_{t-1}+ \phi_2 X_{t-2}+ \phi_3 X_{t-3} +...+ \phi_p X_{t-p}+\theta_0 +\theta_1 W_{t-1} +\theta_2 W_{t-2}+\theta_3 W_{t-3}+...+\theta_q W_{t-q}\]
In the above equation, $X_t$ is the value to predict, $X_{t-1},X_{t-2},X_{t-3}...X_{t-p}$ are past values, $\phi_{p}$ are the coefficients of the past values, $W_{t-1},W_{t-2},W_{t-3}...W_{t-p}$ are past white noise errors, $\theta_{p}$ are the coefficients of the past errors.By fitting this model to oour data, We found that the price of dogecoin fits a ARIMA(2,1,1) model, and by forecating based on this model, the price of dogecoin after 23rd April 2022 will first slightly increase for one day, then decrease for nine days, the overall price of the dogecoin in the next ten days presents a downward trend. 

# Data

In time series analysis, the model created cannot be accurate if the sample size is too small, thus a dataset with over 100 observations is desired for modeling.
Our dataset was collected from Yahoo finance, and it recorded the open price, the high price, the low price, the close price, the volume, and the adjusted price of dogecoin from 1st May 2021 to 23rd April 2022, and thus there are 358 observations. Since there are 358 observations in the dataset, we could build a relatively precise model if there are some underlying relationship between observations. The dogecoin was orignally released in 2013, however, we are not going to use all the observations for modeling, because the price of the dogecoin was so low for a period of time, as Figure \@ref(fig:dogewhole) shown, which may largely influence the accuracy of the model, thus we focus on the price of dogecoin from 1st May 2021 to 23rd April 2022. Based on Figure \@ref(fig:all), observed although all five different prices have different values, there trends and shapes are very similar, thus we only have to anlyze one of these price for modeling, in this case, we choose the closing price. Figure \@ref(fig:volume) shows the volume of the Dogecoin for the past 358 days, although volume could also give investors some insights of whether entering the market or not, in this paper will only focuse on the price.

```{r dogewhole, results=FALSE, echo=FALSE, message = FALSE, fig.cap="The close price of Dogecoin from 2018 to 2022",fig.dim=c(7,4)}
library(quantmod)
library(dplyr)
library(janitor)
library(ggplot2)
library(ggthemes)
getSymbols('DOGE-USD',from="2018-01-01", to="2022-04-23")
doge_whole <- `DOGE-USD`
doge_whole <- doge_whole[,4]
doge_whole <- data.frame(close_price = doge_whole$`DOGE-USD.Close`, Date = time(doge_whole))
doge_whole <- doge_whole %>% clean_names()
ggplot(doge_whole,aes(x=date,y=doge_usd_close)) + 
  geom_line(col="seagreen3") +
  labs(title = "The close price of Dogecoin from January 2018 to April 2022",
       x = "Date",
       y = "Price(USD)") +
  theme(plot.title = element_text(face="bold", hjust = 0.5,  family="serif"),
        axis.title.x = element_text(family="serif"),
        axis.title.y = element_text(family="serif"))

```

```{r all, results=FALSE,echo=FALSE,message=FALSE,fig.cap="Five different prices of the Dogecoin from May 2021 to April 2022",fig.dim=c(7,4)}

getSymbols('DOGE-USD', from="2021-05-01", to="2022-04-23")
doge_all <- `DOGE-USD`
doge_all <- data.frame(doge_close = doge_all$`DOGE-USD.Close`,
                       doge_open = doge_all$`DOGE-USD.Open`,
                       doge_high = doge_all$`DOGE-USD.High`,
                       doge_low = doge_all$`DOGE-USD.Low`,
                       doge_adjusted = doge_all$`DOGE-USD.Adjusted`,
                       doge_volume = doge_all$`DOGE-USD.Volume`,
                       date = time(`DOGE-USD`))
doge_all <- doge_all %>% clean_names()

ggplot() + 
  geom_line(doge_all,mapping=aes(x=date, y=doge_usd_close, color="Opening price")) +
  geom_line(doge_all,mapping=aes(x=date, y=doge_usd_open, color="Closing price")) +
  geom_line(doge_all,mapping=aes(x=date, y=doge_usd_high, color="High price")) +
  geom_line(doge_all,mapping=aes(x=date, y=doge_usd_low, color="Low price")) +
  geom_line(doge_all,mapping=aes(x=date, y=doge_usd_adjusted, color="Adjusted price")) +
  labs(title = "Five different prices from January 2021 to present",
       x = "Date",
       y = "Price(USD)") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold",family="serif"),
        legend.title = element_blank(),
        legend.position = "bottom",
        axis.title.x = element_text(family="serif"),
        axis.title.y = element_text(family="serif"))
```

```{r volume,results=FALSE,echo=FALSE,message=FALSE,fig.cap= "The volume of the Dogecoin from May 2021 to April 2022",fig.dim=c(7,4)}

ggplot(doge_all, aes(x=date,y=doge_usd_volume)) + 
  geom_line(col="darksalmon") +
  labs(title="The volume of the Dogecoin from May 2021 to April 2022",
       x = "Date",
       y = "Volume") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold",family="serif"),
        axis.title.x = element_text(family="serif"),
        axis.title.y = element_text(family="serif"))

```



